---
title: "R_bootcamp_day2_exercises"
author: "Donnelly West"
date: "September 15, 2015"
output: html_document
---
#loading data

#check working directory
getwd()

#set / change working directory
setwd("~/Desktop/R_bootcamp/")
#remember:
# setwd("/") takes you forward (use tab-complete)
# setwd("..") takes you backward one folder

flights = readRDS("flights.RDS")

str(flights)

(more info:http://stat-computing.org/dataexpo/2009/the-data.html)
Get a sense for what the size of the data and what it represents using str, summary, and/or head: What spatial and temporal domain do they cover?

range(flights$year)
2012-2013

range(flights$month)
Jan-Dec

1.How many flights are there in the dataset?
```{r}
nrow(flights)
```

2.Which airline has the most flights? The least?
```{r}
flights %>% 
  count(carrier) %>% 
  arrange(n) %>% 
  slice(c(1, 16))

```

3. How many airlines fly from LAX to SFO?

```{r}
flights$origin

CA = flights %>% 
  group_by(carrier) %>% 
  filter(origin == "LAX" & dest == "SFO") 
#this works..
length(unique(CA$carrier)) #if you're trying to use count in dplyr, it wants the data set separately count(CA, unique(carrier)) - but it still doesn't return the # of carriers, it just returns a neat list of the 6
  
```

4. a.To what airports can you fly from Sacramento (SMF)?
```{r}
SMF = flights %>% 
  filter(origin == "SMF") %>% 
  group_by(dest) 

print(unique(SMF$dest))
#maybe
```

```{r}
SMF2 = flights %>% 
  select(origin, dest) %>% 
  filter(origin == "SMF")
  


print(unique(SMF2$dest))
#who knows, but we got the same answer
```

Michael's solution in class:
```{r}
flights %>% 
  select(origin, dest) %>% 
  filter(origin == "SMF") %>% 
  count(dest) %>% 
  as.data.frame()

```

b.Which of these represents the longest distance flight?
```{r}
flights %>% 
  select(origin, dest, distance) %>% 
  filter(origin == "SMF") %>% 
  arrange(desc(distance)) %>% 
  slice(1)
```


c.Plot the average time to each airport from SMF. Make sure the order of the destination airports makes the plot easy to read
```{r}
# flights %>% 
#   select(origin, dest, air_time) %>% 
#   filter(origin == "SMF") %>% 
#   group_by(dest) %>%
#   summarise(avgTime = mean(air_time)) #get a lot of NA values

flights %>% 
  select(origin, dest, air_time) %>% 
  filter(origin == "SMF") %>% 
  group_by(dest) %>%
  summarise(avgTime = mean(air_time, na.rm=TRUE)) %>% 
  ggplot(aes(x = reorder(dest, avgTime), y = avgTime))+
  geom_point()+
  theme(axis.text.x=element_text(angle=-45,hjust=0))+
  ggtitle("Average Flight Times from Sacramento")+
  ylab("Flight Times (mins)")+
  xlab("Destination Airports")
  
head(flights)
# 
# 
# 
# # flights %>% 
# #   select(origin, dest, air_time) %>% 
# #   filter(origin == "SMF") %>% 
# #   group_by(dest) %>%
# #   summarise(avgTime = mean(.,air_time, na.rm=TRUE)) #worse
# 
# flights %>% 
#   select(origin, dest, air_time) %>% 
#   filter(origin == "SMF") %>% 
#   group_by(dest) %>%
#   summarise(avgTime = mean(air_time, na.omit=TRUE)) #same as first
```

5.a. Identify the ten busiest airports and five highest-volume airlines.
```{r}
# flights %>% 
#   select(carrier, origin, dest) %>%
#   group_by(carrier) %>% 
#   arrange(origin) %>% 
#   as.data.frame()
#    
#   arrange(desc(count(flights, origin))) #error: cannot handle. wow. never seen that before

# airlines
flights %>% 
  count(carrier) %>% 
  top_n(5,n) %>% 
  arrange(-n)

# airports
flights %>% 
  count(c(origin, dest)) %>% 
  top_n(10,n) %>% 
  arrange(-n)
#can't do that either... but ONLY looking at origin OR dest seems limited. need to look at both???
O = flights %>% 
  count(origin) %>% 
  top_n(20,n) %>% 
  arrange(-n)

D = flights %>% 
  count(dest) %>% 
  top_n(20,n) %>% 
  arrange(-n)

class(D)
#how do i compare data frames?

head(O)
head(D)  
#help from Moe
dnew <- cbind(O, D)
dnew
# the top 20 are the same airports, so maybe my concern for dest/origin bias was dumb.
#Moe's code:
D1 <- D %>% 
  select(., airport = dest, dests = n)
O1 <- O %>% 
  select(., airport = origin, origins = n )
head(D1)
head(O1)

?merge

together <- merge(D1, O1, by = 'airport')

head(together)

together %>% 
  mutate(sum = dests + origins) %>% 
  arrange(-sum) %>% 
  top_n(10, n) #finally!

```

b.Among those, compute the number of flights each airline had at each origin airport.

head(flights)
