---
title: "Day2"
author: "Donnelly West"
date: "September 15, 2015"
output: html_document
---

```{r}
summary(cars)
```

#You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

```{r, echo=FALSE}
str(mtcars)
```
#prints all values for this column
mtcars$mpg
#tells you what R thinks the data type is
class(mtcars$wt)
#more info - mine looks different than his..
is(mtcars$wt)

#square bracet subsetting
head(mtcars)
#give me first row, first column 

mtcars[1,1] #it just says "21"?

#get row names
carRownames= rownames(mtcars)
#what class did it make it?
class(carRownames)
#pull first car row name
carRownames[1]
#last row name
#can count
carRownames[32]
#or can be clever
carRownames[length(carRownames)]
#tada

#Create a column on mtcars called "carname" from the rowname

#pull a column out = mtcars$wt, same principle to add a column
head(mtcars)
mtcars$carName = rownames(mtcars)
head(mtcars)

#functional programming - things don't change when you do a function to them, only when you assign a variable to it. you have to use an '=' or nothing will change.

#renaming columns
# This extracts the last column name
colnames(mtcars)[length(colnames(mtcars))]

colnames(mtcars)[length(colnames(mtcars))] = "theCarName"
colnames(mtcars)


#quiz: what does [3,2] return on mtcars?
mtcars[3,2]

#vectorization -> the unit of analysis is never A number, it's a vector or string of numbers. the order is preserved. demonstration w/ independent vectors

x = c(4,7,11)
#so x is a numeric vector, length 3

y= 1:3 #colon is inclusive

#x + y
x+y

x+1 #will add 1 to each value of x
#what if you only want to add 1 to the first?
x[1]+1

?seq()
#so if you type seq(,5,3) it will use 'from = 1' by default
4:12
seq(from=4, to=12, by=4)

#or it will calculate the gap
seq(from=4, to=12, length.out=3) #so will pick first, last, and exact middle

#equivalent to seq(... by =)
seq(4,12,4)

#pop quiz
seq(to=12, length.out=3)
seq(from=1,to=12,length.out=3)
seq(to=12, by=1, length.out=3)
#Michael: it doesn't matter; I don't know what it's doing. the point is you should name your arguements!
#but what's happening is that by=default >> from = default

#random # generator
?sample
sample(1:10, 1)
#is it really random?
sample(1:34694570, 5)

#looks like it

#default state is "replace=false" so you'll never pull the same # twice

#have your 10 samples randomly assorted:
sample(1:10,10)
#doesn't just work with numbers...

letters
sample(letters,3)
#all
sample(letters,length(letters))

#for reproducibility, you can set your seed
set.seed(95616)

#does R studio reset the seed upon close/open? yes!

#for decimales random uniform
runif(1, min=0, max = 1)

#random number from normal distribution
Mynorm = rnorm(1e6)
#1e6 is 1x10^6, default = mean=0, stdev =1
#plot the density quickly using R graphics
plot(density(Mynorm))

#add a column to data.frame
#p2w (power to weight)
#try to get R to tell you the name of the car with the highest p2wt ratio

head(mtcars)

mtcars$p2w = (mtcars$hp/mtcars$wt)
head(mtcars) #victory


?max #for just value
max(mtcars$p2w)

?which #nope
mtcars[max(mtcars$p2w, na.rm=TRUE),] #nope; dunno why though

which.max(mtcars$p2w) #tells you the # of the car = 31
#use that info to pull the name
mtcars$carName[which.max(mtcars$p2w)] #works


#pull out the row with the highest p2w
mtcars[mtcars$p2w == max(mtcars$p2w), ] #works

#Michael's version:
mtcars[which.max(mtcars$p2w), c("carName","p2w")]

#is there a way to know the # of the columns without looking at str??
colnames(mtcars) #gives ALL column names
colnames(mtcars) == "hp" #asking it to find the one that matches hp
which(colnames(mtcars) == "hp") # then 'which one matches' = 4

#want to know total # rows/columns?
dim(mtcars)

nrow(mtcars)
ncol(mtcars)

#adding a column T/F if weight >3
mtcars$bigCar = ifelse(mtcars$wt>3, TRUE, FALSE)
head(mtcars)

#Data types:
class(mtcars$wt)
#numeric
class(mtcars$bigCar)
#logical -most basic variable type

#Coersion happens (T/F is be default 1/0 in most programming languages)
as.numeric(mtcars$bigCar)

#Data types:
#logicals = TRUE or FALSE (TRUE != True, true ; TRUE= T)
